---
title: "covid 19"
author: "Paulo Poblete"
date: "Sept,29,2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
```

The plot above shows that the amount of people alive was bigger in comparison to the people that died. About 3200 people were alive and a couple of hundred were dead. Looking at the  alive plot it shows that there were a lot of women more than men in that group. There were a more dead males in comparison to dead women.  


```{r}
barchartGC(~result + Sex,data=kRdata, type="percent")

```


The percentage plot shows that the alive group there were around 58% were female and around 42% were male. In the dead group there were 62% male and 38% female. In a quick glance it looks like there is twice as much dead males than there is females. 


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex,data=kRdata)
rowPerc(table1)
colPerc(table1)
```

We now have the correct percentages, in the alive group 56% were F and 44% were M 
in the dead group 36% were F and 64% were M. 

## Inferential results 





## Conclusions

