---
title: "covid 19"
author: "Paulo Poblete"
date: "Sept,29,2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```


the first plot above, shows that the amount of people are alive was bigger in comparison to the people that did die. 

the 2nd plot shows that the females group outnumber the names in the alive graph. On the 2nd one it shows the number of the dead, which portrays that there is twice as much dead males in comparison to females 


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex,data=kRdata)
rowPerc(table1)
colPerc(table1)
```

## Inferential results 





## Conclusions

